---
title: "NFL PBP Analytics"
output:
  html_document:
    date: 2/2/2022
    theme: journal
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

Drew Weber

### --------------------------------------------------

```{css, echo = FALSE}

table, td, th {
  border: none;
  padding-left: 1em;
  padding-right: 1em;
  margin-left: auto;
  margin-right: auto;
  margin-top: 1em;
  margin-bottom: 1em;
}

```

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(nflscrapR)
library(devtools)
library(dplyr)
library(nflfastR)
library(stargazer)
library(ggplot2)
library(shiny)
library(na.tools)
library(ggimage)
library(nflreadr)
library(ggrepel)
library(scales)
```


```{r, echo=FALSE, include=FALSE}
#2021 Data
pbp <- load_pbp(2021)
write.csv(pbp, "/Users/drew/Desktop/pbp.csv")
PBP_Rolling_Data <- read.csv("/Users/drew/Desktop/pbp.csv")
pbp2021 <- PBP_Rolling_Data %>% 
	filter(!is_na(epa), play_type=="no_play" | play_type=="pass" | play_type=="run")

pbp2021 <- pbp2021 %>%
	mutate(
	pass = if_else(str_detect(desc, "( pass)|(sacked)|(scramble)"), 1, 0),
	rush = if_else(str_detect(desc, "(left end)|(left tackle)|(left guard)|(up the middle)|(right guard)|(right tackle)|(right end)") & pass == 0, 1, 0),
	success = ifelse(epa>0, 1 , 0)
	) 
```

```{r, echo=FALSE, include=FALSE}
#2020 Data
pbp <- load_pbp(2020)
write.csv(pbp, "/Users/drew/Desktop/202020.csv")
pbp2020 <- read.csv("/Users/drew/Desktop/202020.csv")

depth <- load_rosters(2021)
write.csv(depth, "/Users/drew/Documents/depth.csv")
depth <- read.csv("/Users/drew/Documents/depth.csv")
schedules <- load_schedules(2021)
```

```{r, echo=FALSE, include=FALSE}
#Stats 2021
stats <- load_player_stats(seasons = 2021)
write.csv(stats, "/Users/drew/Desktop/stats.csv")
stats_2021 <- read.csv("/Users/drew/Desktop/stats.csv")
```

```{r, echo=FALSE, include=FALSE}
#Next Gen Stats
nxt_gen_2021_receiving <- load_nextgen_stats(2021, stat_type = "receiving")
write.csv(nxt_gen_2021_receiving, "/Users/drew/Desktop/nxt_gen_2021_receiving.csv")
WR_nxt <- read.csv("/Users/drew/Desktop/nxt_gen_2021_receiving.csv")
WR_nxt <- WR_nxt %>%
  filter(!is_na(avg_cushion), !is_na(avg_separation), !is_na(avg_yac_above_expectation))

nxt_gen_2021 <- load_nextgen_stats(2021)
write.csv(nxt_gen_2021, "/Users/drew/Desktop/nxt_gen_2021.csv")
nxt_2021 <- read.csv("/Users/drew/Desktop/nxt_gen_2021.csv")
```

```{r, include=FALSE, echo=FALSE}
#pbp <- load_pbp(2019)
#write.csv(pbp, "/Users/drew/Documents/Season_19.csv")
pbp2019 <- read.csv("/Users/drew/Documents/Season_19.csv")

#pbp <- load_pbp(2018)
#write.csv(pbp, "/Users/drew/Documents/Season_18.csv")
pbp2018 <- read.csv("/Users/drew/Documents/Season_18.csv")

#pbp <- load_pbp(2017)
#write.csv(pbp, "/Users/drew/Documents/Season_17.csv")
pbp2017 <- read.csv("/Users/drew/Documents/Season_17.csv")
```

```{r, include=FALSE, echo=FALSE}
#nxt_2020 <- load_nextgen_stats(2020)
#write.csv(nxt_2020, "/Users/drew/Documents/nxt_gen_2020.csv")
nxt_2020 <- read.csv("/Users/drew/Documents/nxt_gen_2020.csv")
```

```{r, include=FALSE, echo=FALSE}
nfl_logos_df <- read_csv("https://raw.githubusercontent.com/statsbylopez/BlogPosts/master/nfl_teamlogos.csv")
```

```{r, include=FALSE, echo=FALSE}
qb_20 <- pbp2020 %>%
  filter(!is.na(epa)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(epa),
    cpoe = mean(cpoe, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    team = last(posteam)
  ) %>%
  filter(n_dropbacks > 100 & n_plays > 100)

qb_20 <- qb_20 %>%
  rename( EPA_2020 = epa,
          CPOE_2020 = cpoe,
          n_dropbacks_2020 = n_dropbacks,
          n_plays_2020 = n_plays,
            team_2020 = team)

qb_19 <- pbp2019 %>%
  filter(!is.na(epa)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(epa),
    cpoe = mean(cpoe, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    team = last(posteam)
  ) %>%
  filter(n_dropbacks > 100 & n_plays > 100)

qb_19 <- qb_19 %>%
  rename( EPA_2019 = epa,
          CPOE_2019 = cpoe,
          n_dropbacks_2019 = n_dropbacks,
          n_plays_2019 = n_plays,
            team_2019 = team)

qb_18 <- pbp2018 %>%
  filter(!is.na(epa)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(epa),
    cpoe = mean(cpoe, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    team = last(posteam)
  ) %>%
  filter(n_dropbacks > 100 & n_plays > 100)

qb_18 <- qb_18 %>%
  rename( EPA_2018 = epa,
          CPOE_2018 = cpoe,
          n_dropbacks_2018 = n_dropbacks,
          n_plays_2018 = n_plays,
            team_2018 = team)

qb_17 <- pbp2017 %>%
  filter(!is.na(epa)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(epa),
    cpoe = mean(cpoe, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    team = last(posteam)
  ) %>%
  filter(n_dropbacks > 100 & n_plays > 100)

qb_17 <- qb_17 %>%
  rename( EPA_2017 = epa,
          CPOE_2017 = cpoe,
          n_dropbacks_2017 = n_dropbacks,
          n_plays_2017 = n_plays,
            team_2017 = team)

qb_21 <- pbp2021 %>%
  filter(!is.na(epa)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(epa),
    cpoe = mean(cpoe, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    scram_rate = mean(qb_scramble),
    scram_tot = sum(qb_scramble),
    team = last(posteam)) %>%
  filter(n_dropbacks > 199 & n_plays > 199) %>%
  rename(EPA_2021 = epa,
          CPOE_2021 = cpoe,
          n_dropbacks_2021 = n_dropbacks,
          n_plays_2021 = n_plays,
            team_2021 = team) %>%
  left_join(qb_20, by = c('id', 'name')) %>%
  left_join(qb_19, by = c('id', 'name')) %>%
  left_join(qb_18, by = c('id', 'name')) %>%
  left_join(qb_17, by = c('id', 'name')) %>%
  left_join(teams_colors_logos, by = c('team_2021' = 'team_abbr')) %>%
   filter(!is_na(EPA_2020 | EPA_2019 | EPA_2018)) %>%
  filter(!is_na(EPA_2020)) %>%
  mutate(epa_diff = round(EPA_2021 - EPA_2020, digits = 4)) %>%
  mutate(change_team_2021 = ifelse(team_2021==team_2020, 0,1),
          change_change_2021 = ifelse(team_2021==team_2019, 0,1),
         
         change_team_2020 = ifelse(team_2020==team_2019, 0,1),
         change_change_2020 = ifelse(team_2020==team_2018, 0,1),
         
        change_team_2019 = ifelse(team_2019==team_2018, 0,1),
        change_change_2020 = ifelse(team_2019==team_2017, 0,1))



qb_21_graph <- qb_21 %>%
  filter(!is_na(EPA_2020)) %>%
  mutate(epa_diff = round(EPA_2021 - EPA_2020, digits = 4)) %>%
  summarise(EPA_2021,
            EPA_2020,
            name,
            n_plays_2021,
            team_color,
            epa_diff)
```

## 1. WR Data

```{r, include=FALSE, message=FALSE, warning=FALSE}
#separation <- WR_nxt %>%
 # filter(week==0) %>%
  #group_by(player_short_name) %>%
  #summarize(
   # avg_cushion, 
    #avg_separation,
  #  avg_yac_above_expectation,
   # avg_intended_air_yards,
  #  targets,
   # team = last(team_abbr)
  #) %>%
  #filter(targets > 99) 



separation <- read.csv("/Users/drew/Documents/separation.csv")

separation <- separation %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
separation %>%
  ggplot(aes(x = avg_separation, y = avg_yac_above_expectation)) +

  geom_hline(yintercept = mean(separation$avg_yac_above_expectation), color = "red", linetype = "dashed", alpha=0.5) +

  geom_vline(xintercept =  mean(separation$avg_separation), color = "red", linetype = "dashed", alpha=0.5) +
  
  geom_point(color = separation$team_color, cex=separation$targets / 30, alpha = .6) +

  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +

  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Avg Separation",
       y = "Average YAC Over Expected",
       title = "Graph 1: Individual WR Separation and YAC",
       caption = "Min. 100 Targets") +
 
  theme_bw() +

  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +

  geom_label(aes(x = 3.75, y = -1, label = "Gets Open"), size = 2.5) +
  geom_label(aes(x = 2.2, y = -0.5, label = "Covered up"),  color = "red", size = 2.5) +
  geom_label(aes(x = 2.2, y = 3.5, label = "YAC Guys"),  size = 2.5) +
  geom_label(aes(x = 3.75, y = 3.5, label = "Running Wild"), color = "dark green", size = 2.5) 



```

This graphic shows **Average Separation** and **Average Yards After Catch Over Expected.** Average YAC Over Expected is a metric that measures how receivers do after the catch compared to how they are supposed to.  These are two important metrics on receiver performance. Separation may be the most important thing a WR can generate in football. Combining that with how well they do with the ball in their hands can paint a decent picture of how they are performing. It can be seen that Deebo Samuel is in a world of his own this year. His separation may be somewhat accredited to Shanahan route designs, but it cannot be understated how well he was as a playmaker this season.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
separation %>%
  ggplot(aes(x = avg_separation, y = avg_intended_air_yards)) +
  #horizontal line with mean throw time
  geom_hline(yintercept = mean(separation$avg_intended_air_yards), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(separation$avg_separation), color = "red", linetype = "dashed", alpha=0.5) +

  geom_point(color = separation$team_color, cex=separation$targets / 30, alpha = .6) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Avg Separation",
       y = "Average Intended Air Yards",
       title = "Graph 2: Individual WR Performance",
       caption = "Min. 100 Targets") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) 


```

Separation is not created equal.  It is harder to get open farther down the the field as opposed to catching a bubble screen at the line of scrimmage.  Tyler Lockett gets open extremely deep, unlike how the other receievers are producing.  This will be interesting to see with a new Quarterback in Seattle this year.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
separation %>%
  ggplot(aes(x = avg_cushion, y = avg_separation)) +
  #horizontal line with mean throw time
  geom_hline(yintercept = mean(separation$avg_separation), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(separation$avg_cushion), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = separation$team_color, cex=separation$targets / 25, alpha = .6) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Avg Cushion",
       y = "Average Depth of Target",
       title = "Graph 3: WR Separation",
       caption = "Min 100 Targets") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) 
```

Receivers have an easier time getting open when they are given cushions off the line of scrimmage.  This makes sense as even if a Cornerback is beaten press coverage, they are most likely manned up with the receiver and will not cede much separation.


```{r, include=FALSE, echo=FALSE}
nfl_logos_df <- read_csv("https://raw.githubusercontent.com/statsbylopez/BlogPosts/master/nfl_teamlogos.csv")



separation2 <- WR_nxt %>% 
   filter(week>0) %>%
  group_by(team_abbr) %>%
  summarize(
    mean_cush = mean(avg_cushion), 
    mean_sep = mean(avg_separation),
    mean_eyac = mean(avg_yac_above_expectation),
    mean_targets = mean(targets)
  )  %>%
  left_join(nfl_logos_df, by = c("team_abbr" = "team_code")) %>%
  left_join(teams_colors_logos, by = 'team_abbr')
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
separation2 %>%
  ggplot(aes(x = mean_cush, y = mean_sep, image = team_logo_wikipedia)) +
  #horizontal line 
  geom_hline(yintercept = mean(separation2$mean_sep), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line 
  geom_vline(xintercept =  mean(separation2$mean_cush), color = "red", linetype = "dashed", alpha=0.5) +
  
	geom_image(size = 0.06, by="width") +

 stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Avg Cushion",
       y = "Average Separation",
       title = "Graph 4: Team WR Separation") +
  #bw theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) 
```

This graphic shows Team Separation and Cushion data. This is important because it can show how teams are being played defensively and how they perform against that.


## 2. QB aDot

```{r, include=FALSE, message=FALSE}
#There is an error when loading in next gen stats.  Week 0 for some QBs is NA.  This makes the data visualization incorrect.  I downloaded this as a csv, then edited the NAs.

#time_to_throw_vs_Sticks <- nxt_2021 %>%
 # filter(week==0) %>%
  #group_by(player_gsis_id, player_short_name) %>%
  #mutate(team=last(team_abbr)) %>%
  #summarize(
   # avg_time_to_throw, 
    #avg_air_yards_to_sticks,
    #aggressiveness,
    #completion_percentage_above_expectation,
    #attempts,
    #team
  #) %>%
  #filter(attempts > 299) 

time_to_throw_vs_Sticks <- read.csv("/Users/drew/Documents/time_to_throw.csv")


time_to_throw_vs_Sticks <- time_to_throw_vs_Sticks %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
time_to_throw_vs_Sticks %>%
  ggplot(aes(x = avg_time_to_throw, y = avg_air_yards_to_sticks)) +

  geom_hline(yintercept = mean(time_to_throw_vs_Sticks$avg_air_yards_to_sticks), color = "red", linetype = "dashed", alpha=0.5) +

  geom_vline(xintercept =  mean(time_to_throw_vs_Sticks$avg_time_to_throw), color = "red", linetype = "dashed", alpha=0.5) +

  geom_point(color = time_to_throw_vs_Sticks$team_color, cex=time_to_throw_vs_Sticks$attempts / 100, alpha = .6) +
  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +

  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  labs(x = "Average Time to Throw",
       y = "Average Intended Air Yards from 1st Down",
       title = "Graph 5: 2021 Time to Throw to Sticks",
       caption = "Y=0 is First Down Marker") +

  theme_bw() +

  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = )) +
  geom_label(aes(x = 2.95, y = -2.5, label = "Hold Long, Shallow"), color = "red", size = 2.5) +
  geom_label(aes(x = 2.42, y = -2.5, label = "Cowards"), size = 2.5) +
  geom_label(aes(x = 2.42, y = .9, label = "Quick, Deep"), color = "dark green", size = 2.5) +
  geom_label(aes(x = 2.95, y = .9, label = "Hero Ballers"), size = 2.5) +
  ylim(-2.75, 1.15) +
  xlim(2.38, 3.0)
```

This graphic shows the [**Average Time to Throw**]{.ul} to the [**Average Intended Air Yards From the Sticks**]{.ul}. This means that Y=0 is the first down. This is meant to show how QBs are getting the ball out. The upper left is the best place to be. It means these passers are getting the ball out quickly AND farther than average. Conversely, the bottom right is a bad place to be. It means these passers hang on to the ball long AND throw it shorter than average.

```{r, include=FALSE}
time_to_throw_vs_Sticks <- read.csv("/Users/drew/Documents/time_to_throw.csv")

time_to_throw_vs_Sticks <- time_to_throw_vs_Sticks %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))

time_to_throw_2020 <- nxt_2020 %>%
  filter(week==0) %>%
  group_by(player_gsis_id, player_short_name) %>%
  mutate(team=last(team_abbr)) %>%
  summarize(
   avg_time_to_throw, 
    avg_air_yards_to_sticks,
    aggressiveness,
    completion_percentage_above_expectation,
    attempts
  ) %>%
    filter(attempts>299) 


time_to_throw_2020 <- time_to_throw_2020 %>%
  rename(avg_time_to_throw_2020 = avg_time_to_throw, 
    avg_air_yards_to_sticks_202 = avg_air_yards_to_sticks,
    aggressiveness_2020 = aggressiveness,
    completion_percentage_above_expectation_2020 = completion_percentage_above_expectation,
    attempts_2020 = attempts)


test <- time_to_throw_vs_Sticks %>%
  left_join(qb_21, by = c("player_short_name" = "name")) %>%
  left_join(time_to_throw_2020, by = c("player_short_name" = "player_short_name")) 

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
test_graph2 <- test %>%
  filter(!is_na(avg_air_yards_to_sticks_202))

test_graph2 %>%
  ggplot(aes(x = avg_time_to_throw_2020, y = avg_air_yards_to_sticks_202)) +
 
  geom_hline(yintercept = mean(test_graph2$avg_air_yards_to_sticks_202), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(test_graph2$avg_time_to_throw_2020), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = test_graph2$team_color.y, cex=test_graph2$attempts_2020 / 100, alpha = .6) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm') +
  #titles and caption
  labs(x = "Average Time to Throw",
       y = "Average Air Yards from 1st Down",
       title = "2020 QB Depth of Target",
       caption = "Y=0 is First Down Marker") +
  
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  geom_label(aes(x = 2.95, y = -2.5, label = "Hold Long, Shallow"), color = "red", size = 2.5) +
  geom_label(aes(x = 2.42, y = -2.5, label = "Cowards"), size = 2.5) +
  geom_label(aes(x = 2.42, y = .9, label = "Quick, Deep"), color = "dark green", size = 2.5) +
  geom_label(aes(x = 2.95, y = .9, label = "Hero Ballers"), size = 2.5) 
```

2020 Stats

```{r, echo=FALSE, message=FALSE, warning=FALSE}
time_to_throw_vs_Sticks %>%
  ggplot(aes(x = aggressiveness, y = completion_percentage_above_expectation)) +
  #horizontal line with mean throw time
  geom_hline(yintercept = mean(time_to_throw_vs_Sticks$completion_percentage_above_expectation), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(time_to_throw_vs_Sticks$aggressiveness), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = time_to_throw_vs_Sticks$team_color, cex=time_to_throw_vs_Sticks$attempts / 100, alpha = .6) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=player_short_name), size = 2.75 ) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Aggresiveness Index",
       y = "CPOE",
       title = "Graph 6: Agressiveness Index",
       caption = "Y=0 is First Down Marker") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +

  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = )) 
```




## 3. QB Short Report: How Consistent is Quarterbacking?

How consistent is NFL QB play?  Can we look at previous seasons to predict forward for QB play?  What about when players change teams, such as Matthew Stafford?  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
qb_years_reg_21 <-lm(EPA_2021 ~  EPA_2020 + EPA_2019 + EPA_2018 + EPA_2017, data = qb_21)

qb_years_reg_20 <-lm(EPA_2021 ~ EPA_2020 + EPA_2019 + EPA_2018 +change_team_2019, data = qb_21)

qb_years_reg_19 <-lm(EPA_2021 ~ EPA_2020 + EPA_2019 + change_team_2020, data = qb_21)

qb_years_reg_18 <-lm(EPA_2021 ~ EPA_2020 + change_team_2021, data = qb_21)


qb_cpoe_2017 <-lm(CPOE_2021 ~  CPOE_2020 + CPOE_2019 + CPOE_2018 + CPOE_2017, data = qb_21)

qb_cpoe_2018 <-lm(CPOE_2021 ~ CPOE_2020 + CPOE_2019 + CPOE_2018, data = qb_21)

qb_cpoe_2019 <-lm(CPOE_2021 ~ CPOE_2020 + CPOE_2019, data = qb_21)

qb_cpoe_2020 <-lm(CPOE_2021 ~ CPOE_2020, data = qb_21)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
stargazer(qb_years_reg_21, qb_years_reg_20, qb_years_reg_19, qb_years_reg_18,
          type="html",
          title = "QB EPA over Years",
          style = "aer",
          column.labels = c("Since 2017", "Since 2018", "Since 2019", "Since 2020"),
          covariate.labels = c("2020 EPA", "2019 EPA", "2018 EPA", "2017 EPA", "New Team in 2019", "New Team in 2020", "New Team in 2021"),
          dep.var.labels = c("2021 EPA"))

```

The farther we go back, the better goodness of fit we achieve.  2020 EPA is correlated at a statistically significant level in 3 of the 4 regressions.  We can see that changing teams does not have a statistically significant effect on EPA.  This is very intriguing as it poses the hypothesis that maybe a QB is just who they are.  Something I wish that could be expanded on is the scheme that the QB is in.  Scheme matters heavily in difficulty, comfort, and production for QBs.  Finding a way to apply this to data is something I hope to do in the future, if possible.


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
stargazer(qb_cpoe_2017, qb_cpoe_2018, qb_cpoe_2019, qb_cpoe_2020,
          type="html",
          title = "QB CPOE over Years",
          style = "aer",
          covariate.labels = c("2020 CPOE", "2019 CPOE", "2018 CPOE", "2017 CPOE"),
          column.labels = c("CPOE"),
          dep.var.labels = c("2021 CPOE"))
          
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
qb_21_graph %>%
  ggplot(aes(x = EPA_2020, y = EPA_2021)) +

  geom_hline(yintercept = mean(qb_21_graph$EPA_2021), color = "red", linetype = "dashed", alpha=0.5) +

  geom_vline(xintercept =  mean(qb_21_graph$EPA_2020), color = "red", linetype = "dashed", alpha=0.5) +

  geom_point(color = qb_21_graph$team_color, cex=qb_21_graph$n_plays_2021 / 150, alpha = .6) +

  geom_text_repel(aes(label=name), size = 3.25) +

  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "2020 EPA",
       y = "2021 EPA",
       title = "EPA 2020 & 2021",
       caption = "Min. 200 2021 Dropbacks") +
  
  theme_bw() +
 
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold")
  ) +

  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
```

Simple. Above the reference line, improvement of EPA. Below, a worse EPA.

```{r, warning=FALSE, echo=FALSE, message=FALSE}
head_epa_diff <- qb_21 %>%
  select(name, epa_diff) %>%
  arrange(desc(epa_diff)) %>%
  head() 

tail_epa_diff <- qb_21 %>%
  select(name, epa_diff) %>%
  arrange(desc(epa_diff)) %>%
  tail() 

head_epa_diff <- subset(head_epa_diff, select = -c(id))
tail_epa_diff <- subset(tail_epa_diff, select = -c(id))

```


```{r, results ='asis', echo=FALSE}
stargazer(head_epa_diff, tail_epa_diff,
          summary = FALSE,
          style = "aer",
          type = "html",
          title = "20 to 21 EPA per Dropback Difference")
```

We can see a few of the best and worst of EPA per dropback difference from 2020 to 2021.  It is interesting that Jared Goff and Matt Stafford went different directions of almost the same amount.  Goff and Stafford were traded for each other before 2021.  The Rams added about 0.15 EPA per dropback by doing this.  It may not seem like much, but it is a massive upgrade.  That is over 5 more expected points a game when around 30 dropbacks a game.  Even more impressive is that Stafford was throwing lots of pick sixes (massively, massively negative EPA plays)  and he still put up this difference over Goff!

```{r, echo=FALSE, message=FALSE, warning=FALSE}
test_graph3 <- test %>%
  select(EPA_2021, EPA_2020, player_short_name) %>%
  filter(!is_na(EPA_2020)) %>%
  left_join(qb_21, by = c("player_short_name" = "name")) %>%
  select(EPA_2021.x, EPA_2020.x, player_short_name, epa_diff) %>%
  rename(EPA_2021 = EPA_2021.x,
         EPA_2020 = EPA_2020.x,
         name = player_short_name,
         epa_diff = epa_diff)

test_graph3 %>%
  ggplot(aes(x = name)) +
  
 geom_point(aes(y=EPA_2021, color = "red")) +
  geom_point(aes(y=EPA_2020, color = "black")) +
  # Custom the Y scales:
  scale_y_continuous(
    name = "EPA 2021",
    sec.axis = sec_axis( trans=~., name="EPA 2020")) + 
  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 

  
ttt_reg <- lm(data = test, avg_time_to_throw ~ avg_time_to_throw_2020)
```

## 4. QB Sack Data

```{r, message=FALSE, include=FALSE}
sack_data <- pbp2021 %>%
  filter(sack==1) %>%
  group_by(id, name) %>% 
  summarize(
    av_epa_lost = round(mean(epa), digits = 2),
    total_epa_lost = round(sum(epa), digits = 2),
    team = last(posteam),
    Sacks = sum(pass)) %>%
  arrange(desc(av_epa_lost)) %>%
    filter(Sacks > 19)


sack_data <- subset(sack_data, select = -c(id))
sack_data <- sack_data %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))

sack_data_table <- pbp2021 %>%
  filter(sack==1) %>%
  group_by(name) %>% 
  summarize(
    av_epa_lost = round(mean(epa), digits = 2),
    total_epa_lost = round(sum(epa), digits = 2),
     team = last(posteam),
    Sacks = sum(pass)) %>%
  arrange(desc(av_epa_lost)) %>%
  filter(Sacks > 19)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sack_data %>%
  ggplot(aes(x = total_epa_lost, y = av_epa_lost)) +
  #Horizontal line mean EPA
  geom_hline(yintercept = mean(sack_data$av_epa_lost), color = "red", linetype = "dashed", alpha=0.5) +
  #Vertical line mean CPOE
  geom_vline(xintercept =  mean(sack_data$total_epa_lost), color = "red", linetype = "dashed", alpha=0.5) +
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = sack_data$team_color, cex=sack_data$Sacks / 10, alpha = .75) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=name), size = 3.2) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Total EPA Lost via Sacks",
       y = "Average EPA Lost via Sacks",
       title = "Graph 6: EPA Lost by Sacks",
       caption = "Min. 12 Sacks") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  #if this doesn't work, `install.packages('scales')`
  scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 8)) +

  geom_label(aes(x = -50, y = -2.5, label = "Fewer but Costly Sacks"), size = 2.5) +
  geom_label(aes(x = -100, y = -2.5, label = "Lots of Bad Sacks"),  color = "red", size = 2.5) +
  geom_label(aes(x = -100, y = -1.4, label = "Lots of Sacks"),  size = 2.5) +
  geom_label(aes(x = -45, y = -1.38, label = "Not Costly"), color = "dark green", size = 2.5) 




```

Not all sacks are created equal. A sack on 1st and 10 is most likely much worse than a sack on 3rd and 20. It can ruin a drive or can be almost the same as an incompletion. This graphic shows how much [EPA is being lost on average and in total]{.ul}**.** The outliers are what should be noted most from this graphic. Wentz is a catastrophe, whereas Burrow lost the most EPA but wasn't killing the Bengals every time he did it. Lamar Jackson and Josh Allen are freaks and it is interesting to see if the sack "avoiders" remain this way year over year.\`

```{r, results ='asis', echo=FALSE}
stargazer(head(sack_data_table, n = 3), tail(sack_data_table, n = 3),
          summary = FALSE,
          style = "aer",
          type = "html",
          title = "EPA Lost From Sacks")
```

## 5. Situational Regressions and Betting Collinearity

```{r, include=FALSE, message=FALSE, warning=FALSE}
EPAreg1 <-lm(epa~  as.factor(down) + rush_attempt+ pass_attempt, data = pbp2021)


EPAreg2 <-lm(epa~  rush_attempt, data = pbp2021, na.rm = TRUE, subset = down==1)


EPAreg3 <-lm(epa~  rush_attempt + pass_attempt, data = pbp2021, na.rm = TRUE, subset = down==4)


EPAreg4 <- lm(epa~ qb_scramble, data = pbp2021)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
stargazer(EPAreg1, EPAreg2, EPAreg3,EPAreg4,
          type="html",
          title = "EPA Regressions",
          style = "aer",
          covariate.labels = c("2nd Down", "3rd Down", "4th Down", "Rush Attempt", "Pass Attempt", "QB Scramble"),
          column.labels = c("Play Type and Down", "First Down", "Fourth Down", "QB Scramble"),
          dep.var.labels = c("EPA"))
          
```

This is the second straight year that QB Scramble has over a 0.4 EPA coefficient. It is something I always thought about as QBs tend to not make mistakes on scrambles and can usually gain a few yards with their legs at the least. Many young QBs play far better out of structure as they can turn their processor off and just ball.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
catch_yards_reg <-lm(yards_gained ~  complete_pass, data = pbp2021)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
stargazer(catch_yards_reg,
          type="html",
          title = "Catch to Yards Correlation",
          style = "aer",
          covariate.labels = c("Yards Gained"),
          column.labels = c("Yards"),
          dep.var.labels = c("Yards Gained"))
          
```

This was purely made from a betting standpoint. Parlaying a receivers catch props with their receiving props can be profitable if

A. The book allows you.

B. [**They don't take into account this colinearity.**]{.ul}

For example, if I bet on a receiver to get 26 yards, they most likely need 2 or more catches to do so. This means if I expect them to get 26 yards, I should throw in the 2 catch prop as well.

The reason this matters is because books are betting that [at least]{.ul} one of the bets won't hit. In this situation, it is most likely that either 0 of 2 hit, or 2 of 2 hit.

This is very dependent on books NOT discounting for this colinearity. If they are, it needs to be analyzed at the discount rate they are using.

## 6. Rushing Data

```{r, message=FALSE, include=FALSE}
EPA_by_Run_Gap_2020 <- pbp2020 %>%
	filter(wp>.20 & wp<.80 & half_seconds_remaining>120) %>%
	group_by(run_gap) %>%
  filter(!is_na(run_gap), play_type=="no_play" | play_type=="pass" | play_type=="run") %>%
	summarize(mean_epa = round(mean(epa), digits = 3), plays=n()) %>%
	arrange(mean_epa) 
  colnames(EPA_by_Run_Gap_2020) <- c("Run Gap", "Mean EPA", "Plays")
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
stargazer(EPA_by_Run_Gap_2020,
          summary = FALSE,
          style = "aer",
          type = "html",
          title = "2020 EPA by Run Gap")
```

```{r, echo=FALSE, message=FALSE, include=FALSE}
EPA_by_Run_Gap <- pbp2021 %>%
	filter(wp>.20 & wp<.80 & half_seconds_remaining>120) %>%
	group_by(run_gap) %>%
  filter(!is_na(run_gap), play_type=="no_play" | play_type=="pass" | play_type=="run") %>%
	summarize(mean_epa = round(mean(epa), digits = 3), plays=n()) %>%
	arrange(mean_epa) 
  colnames(EPA_by_Run_Gap) <- c("Run Gap", "Mean EPA", "Plays")
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
stargazer(EPA_by_Run_Gap,
          summary = FALSE,
          style = "aer",
          type = "html",
          title = "2021 EPA by Run Gap")
```

```{r, include=FALSE, message=FALSE}
success_rate_runs <- pbp2021 %>%
   group_by(posteam) %>%
  filter(!is.na(epa), play_type=="run") %>%
  summarize(
    success_rate = round(mean(success), digits = 2),
    avg_wp = mean(wp),
    ypc = mean(yards_gained)
  ) 

nfl_logos_df <- read_csv("https://raw.githubusercontent.com/statsbylopez/BlogPosts/master/nfl_teamlogos.csv")
success_rate_runs <- success_rate_runs %>% left_join(nfl_logos_df, by = c("posteam" = "team_code"))

```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
success_rate_runs %>%
  ggplot(aes(x = success_rate, y = avg_wp)) +
  #horizontal line with mean throw time
  geom_hline(yintercept = mean(success_rate_runs$avg_wp), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(success_rate_runs$success_rate), color = "red", linetype = "dashed", alpha=0.5) +

	geom_image(aes(image = url), size = 0.06) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Success Rate on Runs",
       y = "Average Win Probability at All Times",
       title = "Graph 9: Rushing Success Rate") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) 

```

Results are most likely spurious as teams who are winning are going to run the ball the most.

## 7. Pure Passing EPA Table

```{r, echo=FALSE, message=FALSE, include=FALSE}
EPA_Leaders_Twenty_One <- pbp2021 %>%
	filter(pass == 1, !is.na(epa)) %>%
  group_by(id, name) %>%
	summarize(team = last(posteam), 
	          mean_epa = round(mean(epa), digits = 2), 
	          success_rate = round(mean(success), digits = 2), 
	          plays=n())%>%
	arrange(desc(mean_epa)) %>%
  filter(plays>300)

colnames(EPA_Leaders_Twenty_One) <- c('ID', 'Passer', 'Team', 'Mean EPA', 'Success Rate', 'Passes')
EPA_Leaders_Twenty_One <- subset(EPA_Leaders_Twenty_One, select = -c(ID))
EPA_Leaders_Twenty_One = EPA_Leaders_Twenty_One[-1,]
```

```{r, results ='asis', echo=FALSE}
stargazer(head(EPA_Leaders_Twenty_One, n = 8),
          summary = FALSE,
          style = "aer",
          type = "html",
          title = "Passing EPA Leaders")

```

## 8. Trick Play Passing Data

```{r, echo=FALSE, message=FALSE, warning=FALSE}
Non_QB_Passers <- pbp2021 %>%
  filter(!is.na(epa), !is.na(cpoe)) %>%
  group_by(id, name) %>%
  summarize(
    epa = mean(qb_epa),
    cpoe = mean(cpoe),
    yards = mean(yards_gained, na.rm = T),
    n_dropbacks = sum(pass),
    n_plays = n(),
    team = last(posteam)
  ) %>%
  filter(n_dropbacks < 5 & n_dropbacks > 0 & n_plays > 0)

Non_QB_Passers <- Non_QB_Passers %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))

Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$yards != "0", ] 

Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "J.Rosen", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name !="M.Trubisky", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "B.Hoyer", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name !="B.Allen", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "J.Wolford", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "J.Flacco", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "C.Beathard", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "G.Minshew II", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "F.Franks", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "K.Mond", ] 
Non_QB_Passers <- Non_QB_Passers[Non_QB_Passers$name != "M.Mariota", ] 

 
#Non_QB_Passers <- Non_QB_Passers %>% 
  #mutate_all(~replace(., is.nan(.), 0))
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}

Non_QB_Passers %>%
  ggplot(aes(x = cpoe, y = epa)) +
  #horizontal line with mean EPA
  geom_hline(yintercept = mean(Non_QB_Passers$epa), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(Non_QB_Passers$cpoe), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = Non_QB_Passers$team_color, cex=Non_QB_Passers$n_dropbacks / .5, alpha = .6) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=name)) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "CPOE",
       y = "EPA per Play",
       title = "Graph 11: Non-QB Passers EPA",
       caption = "Non-Primary QBs") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  #if this doesn't work, `install.packages('scales')`
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))



```

Just for fun!
